\documentclass{article}
\input{header}
\begin{document}
\pagestyle{empty}
\maketitle
\subtitle{Tutorial Week 5: Topics 1, 2 \& 3}

\paragraph{Explanation on accumulating Syntax Skills:} Let me stress that the point of the syntax block is to learn how to think like a syntactician, learn about English, and try to have fun. But marks are involved, and here's how we mark Syntax Skills: generally speaking, it's enough to show them once. This means that if we have questions Q1A and Q2B in a given tutorial sheet (for example), both of which target skill 6b, then getting one of them right is enough. That said, it's best to attempt all questions.

Some tutorial questions are primarily \textbf{for discussion}. This means that they can count towards your skill tally, but don't worry if you don't get them right -- their goal is to allow for extended discussion in the tutorial itself and some will go beyond the discussion in class.

(Please remember that we're developing this approach to tutorials and marking together, so your feedback is welcome -- keep it specific and try to give it directly to Itamar.)

\section*{Question 1}
\hfill{}
%\ref{whatissyntaxA} $\Box$,
\ref{whatissyntaxB} $\Box$ 
%\ref{whatissyntaxC} $\Box$

\paragraph{Q1A} Given the following list of constituents, visually represent the structure of the sentence \emph{My cousin gave the dog a juicy bone}.
\begin{itemize}
    \item my cousin
    \item the dog
    \item a juicy bone
    \item gave the dog a juicy bone
\end{itemize}
Don’t worry too much about whether your representation is \emph{right} or not in other aspects, just as long as it
shows those four strings as constituents. For example, it doesn't need to have any labels for the constituents (like VP or NP).

\hfill{}\ref{constituencytestVP} $\Box$ %, \ref{diagnosticlimits} $\Box$

\paragraph{For discussion} Above, you were told that \emph{gave the dog a juicy bone} was a constituent.
However, if we apply an \emph{it-cleft} test for constituency to this string we get the following results:
\begin{exe}
    \ex[*]{It is gave the dog a juicy bone that my cousin}
\end{exe}
What does this mean for the representation you gave above?
Can you use other diagnostics to resolve this issue (one is enough but if you can give more that's even better)?

\section*{Question 2}
\hfill{}
\ref{constituencytestNP} $\Box$,
\ref{constituencytestVP} $\Box$,
\ref{constituencytestother} $\Box$
%, M3 $\Box$, M4 $\Box$

\paragraph{Q2A }Which of the following trees is the best representation of the structure for the sentence \emph{She placed that vase on the table}, based on what you can determine about the constituent
structure of the sentence? For each of the trees that you didn't choose in (\ref{Q2_trees}), what is the evidence against it?

\begin{exe}
\ex{
\begin{xlist}
    \ex{
    \small \begin{forest}
for tree={s=2mm, l-=5mm},
	[S [NP[Pronoun\\\emph{she}]] [VP [V\\\emph{placed}] [NP [Det\\\emph{that}] [NP [N\\\emph{vase}] [PP [P\\\emph{on}] [NP [Det\\\emph{the}] [N\\\emph{table}]]]]]]]
    \end{forest}}
    \ex{
    \small \begin{forest}
for tree={s=2mm, l-=5mm},
	[S [NP[Pronoun\\\emph{she}]] [VP [V\\\emph{placed}]
    [NP 
        [Det\\\emph{that}]
        [N\\\emph{vase}]
        [P\\\emph{on}]
        [NP [Det\\\emph{the}] [N\\\emph{table}]]]]]
    \end{forest}}
\ex{
    \small \begin{forest}
for tree={s=2mm, l-=5mm},
	[S [NP[Pronoun\\\emph{she}]] [VP [V\\\emph{placed}] [NP [Det\\\emph{that}] [NP [N\\\emph{vase}]]] [PP [P\\\emph{on}] [NP [Det\\\emph{the}] [N\\\emph{table}]]]]]
    \end{forest}}
    \end{xlist}}
    \label{Q2_trees}
\end{exe}

\newpage
\section*{Question 3}
\hfill{}
\ref{syn_sem_roles} $\Box$,
\ref{sem_roles} $\Box$


As you know, some verbs in English can appear either with two syntactic arguments, or just one:
\begin{exe}
    \ex{
    \begin{xlist}
        \ex{This film shocks.\tabto{10em} This film shocks a lot of people.}
        \ex{The donkey kicked.\tabto{10em} The donkey kicked the door.}
        \ex{The canary drank.\tabto{10em} The canary drank some water.}
    \end{xlist}}
    \label{semantic_rolesA}
    \ex{
    \begin{xlist}
        \ex{The branch broke.\tabto{10em} The bear broke the branch.}
        \ex{The cup dropped.\tabto{10em} The child dropped the bag.}
        \ex{The lake thawed.\tabto{10em} The sun thawed the ice.}
    \end{xlist}}
    \label{semantic_rolesB}
\end{exe}

Thinking of the ways in which participant roles (like AGENT or THEME) are mapped onto syntactic functions (like SUBJECT or OBJECT):

\paragraph{Q3A} What is the difference between the verbs in (\ref{semantic_rolesA}) and those in (\ref{semantic_rolesB})? Describe it in terms of the syntactic and semantic arguments of the predicates.

\paragraph{For discussion:} Can you think of other English verbs that behave like the verbs in (\ref{semantic_rolesA})? And like the verbs in (\ref{semantic_rolesB})?

\paragraph{For discussion:} If you speak or are familiar with another language, are there verbs in that language that behave to some extent (or exactly) like the verbs in (\ref{semantic_rolesB})? If you find a pattern that is similar but not identical, can you describe it?

\end{document}