\documentclass{beamer}
\input{header_beamer}
\subtitle{Topic 10: Nonfinite clauses part 2}

\begin{document}
\maketitle

\frame{\frametitle{Recap}
      \begin{block}{Topic 9 skills}
    \begin{itemize}
        \item[9a]{Identify non-finite clauses headed by infinitival \emph{to} in a principled way.}
        \item[9b]{Distinguish raising structures from control structures on the basis of $\theta$-role assignment.}
    \end{itemize}
    \end{block}
}

\frame{

    The plan for our final week:
        \begin{itemize}
            \item Topic 10: more control/raising
            \item Topic 11: passives
            \item Topic 12: \emph{wh}-questions
        \end{itemize}
}

\frame{\frametitle{Topic 10}
   \begin{block}{Topic 10 skills}
    \begin{itemize}
        \item[10a]{Motivate raising to subject position (A-movement) in terms of either Case or the EPP.}
        % \item[10b]{\sout{Analyse instances of object control in a way that is consistent with the framework developed.}}
    \end{itemize}
    \end{block}

    Research Skills:
    \begin{itemize}
    \item[M1]{State a hypothesis clearly}
    \item[M2]{Present data in support of a hypothesis}
    \item[M3]{Define the conditions under which a hypothesis would be refuted}
    \item[M4]{Present data to refute a hypothesis}
    \item[M5]{Consider the limitations of a diagnostic test when interpreting the results}
    \end{itemize}
}

\frame{
    Verb classes for our exercise:
    \begin{enumerate}
    \item Class 1: %subj control
        \emph{try, hope, want, start, stop, \dots{}}
    \item Class 2: %obj control
        \emph{convince, persuade, tell, \dots{}}
    % \item Class 3: %ambig
        % \emph{promise, ask, \dots{}}
    \item Class 3: %ECM
        \emph{believe, consider, want, expect, allow, \dots{}}
    \item Class 4: %raising
        \emph{seem, appear, eager, start, stop, \dots{}}
    \end{enumerate}

    Construct examples and apply the following tests:
    \begin{itemize}
        \item Idioms
        \item Expletive subjects
        \item Agent-oriented adverbs
        \item Passivised subjects
    \end{itemize}
    See if you can draw trees for the examples you came up with. What decisions do you have to make?
        % I believe DP to IP
    % I believe expletive to IP
}

\frame{ \centering \small
\hspace*{-2.5em}
\begin{tabular}{l|c|c|c|c}
    & Subj control & Obj control & ECM & Raising\\\hline
    & \emph{try} & \emph{convince} & \emph{believe} & \emph{seem}\\
    Embedded subj & PRO (Agent) & PRO (Theme) & Raised object & Raised subject\\
\end{tabular}
}

\frame{
    The subject requirement (raising) has received various treatments:
        \begin{itemize}
            \item Case.
            \item EPP requirement on \obar{I}.
        \end{itemize}
}


\frame{\frametitle{Summary}
   \begin{block}{Topic 10 skills}
    \begin{itemize}
        \item[10a]{Motivate raising to subject position (A-movement) in terms of either Case or the EPP.}
        % \item[10b]{\sout{Analyse instances of object control in a way that is consistent with the framework developed}.}
    \end{itemize}
    \end{block}

    We also discussed:
    \begin{itemize}
        \item Differences between Object Control and ECM.
        \item Ambiguous control/raising verbs like \emph{start, stop}.
    \end{itemize}

    Research Skills:
    \begin{itemize} \small
    \item[M1]{State a hypothesis clearly}
    \item[M2]{Present data in support of a hypothesis}
    \item[M3]{Define the conditions under which a hypothesis would be refuted}
    \item[M4]{Present data to refute a hypothesis}
    \item[M5]{Consider the limitations of a diagnostic test when interpreting the results}
    \end{itemize}
}
\end{document}

a. John expected it to be raining.
b. John expected there to be an earthquake.
c. John expected it to be surprising that she was such a fool.
d. John expected the cat to be out of the bag.
e. John expected the shit to hit the fan.
f. John expected tabs to be kept on him by the FBI.
(19) a. * John persuaded it to be raining.
b. * John persuaded there to be an earthquake.
c. * John persuaded it to be surprising that she was such a fool.
d. * John persuaded the cat to be out of the bag.
e. * John persuaded the shit to hit the fan.
f. * John persuaded tabs to be kept on him by the FBI.

OC made

ECM found to be charming